**Selective DNA and Protein Isolation from Marine Macrophyte Surfaces**

### 2020-04-20 submission/manuscript.Rmd
Added chloroplast plot.

An alternative to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer. After the incubation is done the undisrupted host tissue is removed and the isolation procedure continues omitting host material contaminations. To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009. In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011b] and metagenomes sequencing [@Burke2011a]. This method, thought providing a selective isolation procedure, is using in the extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide available and whose composition is not know [@Burke2009]. Also to our knowledge, no selective isolation protocol for proteins from epiphytic communities inhabiting marine macrophytes was established.

@@

### Illumina 16S rRNA sequencing
An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure. In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016]. These primers contained on their 5' end a tagged sequence. Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5 Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield. The column was eluted in 30 \si{\ul} of deionized water. Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany. Before sequencing, at IMGM the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR. In addition, these primers contained adapter and sample-specific index sequences. The second PCR was carried out for 8 cycles. Beside samples, a positive and negative control were sequenced. A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used. The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.

### Sequence analysis
```{r}
# Calculation of the error rate based on the ATCC MSA-1002 mock community
error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
  filter(numparents==1) %>%
  mutate(mismatches_weight=weight*mismatches) %>%
  mutate(total_weight=weight*total)
error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
# Calculation of the number of sequences in the negative control
nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
  filter(taxon=="Root") %>%
  select(NC_1)
```
Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/). For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014]. Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 138) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database. Reference sequences close to imported ones were selected and used to calculate a phylogenetic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006]. Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology. Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**). Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016]. In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.

@@

### Metaproteomics
Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions [@Wisniewski2009] with small modifications. Before loading the solution to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles. The first washing step after protein solution loading was repeated twice. In addition, centrifugation steps were prolonged if the column was clogged. Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}. The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis. Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy approach was used to reduce the probability of false peptide identification. Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept. For protein identification a minimum of of two peptides and one unique peptide were required. For protein quantification, a chromatographic peak area-based free quantitative method was applied.

@@

@@

>>>>

An alternative to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer. After the incubation is done the undisrupted host tissue is removed and the isolation procedure continues omitting host material contaminations. To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009. In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011a] and metagenomes sequencing [@Burke2011]. This method, thought providing a selective isolation procedure, is using in the extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide available and whose composition is not know [@Burke2009]. Also to our knowledge, no selective isolation protocol for proteins from epiphytic communities inhabiting marine macrophytes was established.

@@

### Illumina 16S rRNA sequencing
An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure. In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016]. These primers contained on their 5' end a tagged sequence. Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5 Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield. The column was eluted in 30 \si{\ul} of deionized water. Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany. Before sequencing, at IMGM the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR. In addition, these primers contained adapter and sample-specific index sequences. The second PCR was carried out for 8 cycles. Beside samples, a positive and negative control were sequenced. A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used. Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.

### Sequence analysis
```{r}
# Calculation of the error rate based on the ATCC MSA-1002 mock community
error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
  filter(numparents==1) %>%
  mutate(mismatches_weight=weight*mismatches) %>%
  mutate(total_weight=weight*total)
error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
# Calculation of the number of sequences in the negative control
nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
  filter(taxon=="Root") %>%
  select(NC_1)
```
Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/). For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014]. Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (Ribosomal Database Project; http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014]. In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences. Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020). Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016]. In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.

@@

### Metaproteomics
Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions [@Wisniewski2009] with small modifications. Before loading the solution to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles. The first washing step after protein solution loading was repeated twice. In addition, centrifugation steps were prolonged if the column was clogged. Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}. The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis. Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011 using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy approach was used to reduce the probability of false peptide identification. Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept. For protein identification a minimum of of two peptides and one unique peptide were required. For protein quantification, a chromatographic peak area-based free quantitative method was applied.

@@

**\autoref{chloroplast}.** \nameref{chloroplast}

@@

\newpage
```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of chloroplast sequences from surfaces of two marine macrophytes (\\textit{C. nodosa} and \\textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{chloroplast}", fig.pos="ht"}
include_graphics("../results/figures/chloroplast.jpg")
```
